---
title: "Lab 4 Demo"
date: "2024-04-22"
output: html_document
---

```{r, include = FALSE}
library(tidytext)
library(tidyverse)
library(tidymodels)
library(textrecipes)
library(discrim) # naive-bayes
library(naivebayes)
```

This data set includes more possible predictors than the text alone, but for this model we will only use the text variable
```{r}

urlfile ="https://raw.githubusercontent.com/MaRo406/EDS-231-text-sentiment/main/data/climbing_reports_model_dat.csv"
incidents_df<-readr::read_csv(url(urlfile))
```

Now we'll split our data into training and test portions

```{r}
set.seed(1234)

incidents2class <- incidents_df %>% 
  mutate(fatal = factor(ifelse(is.na(Deadly), "non-fatal", "fatal")))

incidents_split <- initial_split(incidents2class, strata = fatal) # strata addresses class                                                                            imbalance

incidents_train <- training(incidents_split)
incidents_test <- testing(incidents_split)

```

We use recipe() to specify the predictor and outcome variables and the data.

```{r }
incidents_rec <-recipe(fatal ~ Text, data = incidents_train)
```

Next we add some familiar pre-processing steps on our Text variable: tokenize to word level, filter to the most common words, and calculate tf-idf.

```{r }
recipe <- incidents_rec %>% 
  step_tokenize(Text) %>% 
  step_tokenfilter(Text, max_tokens = 1000) %>% 
  step_tfidf(Text)
```

Create  tidymodels workflow to combine the modeling components

```{r }
incidents_wf <- workflow() %>% 
  add_recipe(recipe) 
```

```{r }
nb_spec <- naive_Bayes() %>% 
  set_mode("classification") %>% 
  set_engine("naivebayes")
```

Now we are ready to add our model to the workflow and fit it to the training data

```{r }
nb_fit <- incidents_wf %>% 
  add_model(nb_spec) %>% 
  fit(data = incidents_train)
```
Next up is model evaluation. We'll stretch our training data further and use resampling to evaluate our naive Bayes model. Here we create 10-fold cross-validation sets, and use them to estimate performance.

```{r }
set.seed(650)
incidents_folds <- vfold_cv(incidents_train)
```

```{r }
nb_wf <- workflow() %>% 
  add_recipe(recipe) %>% 
  add_model(nb_spec)
```

To estimate its performance, we fit the model to each of the resampled folds, and then evaluate on the heldout part of each resampled fold.

```{r }
nb_rs <- fit_resamples(
  nb_wf,
  incidents_folds,
  control = control_resamples(save_pred=T))
```

Extract the relevant information using collect_metrics() and collect_predictions() and examine the performance metrics.

```{r }
nb_rs_metrics <- collect_metrics(nb_rs)

nb_rs_predictions <- collect_predictions(nb_rs)
```

We'll use two performance metrics: accuracy and ROC AUC.
Accuracy is the proportion of the data that is predicted correctly. 
The ROC curve plots the true positive rate against the false positive rate; AUC closer to 1 indicates a better-performing model, while AUC closer to 0.5 indicates a model that does no better than random guessing.

```{r }
nb_rs_predictions %>% 
  group_by(id) %>% 
  roc_curve(truth=fatal, .pred_fatal) %>% 
  autoplot() +
  labs(title="ROC curve for climbing incidence reports")
```

Another model method involves the confusion matrix. A confusion matrix tabulates a model's false positives and false negatives for each class.

```{r }
conf_mat_resampled(nb_rs, tidy=F) %>% 
  autoplot(type="heatmap")
```

